### 2013-07-01 doc/dissertation/document/chapters/parallel_mc/parallel_mc.tex
adding efficiency plots to scaling section. updating multiple set weak scaling definition

%%---------------------------------------------------------------------------%%
\section{Domain Decomposed Neumann-Ulam Algorithm\ }
\label{sec:asynchronous_algorithm}
In the context of radiation transport, in 2009 Brunner and colleagues
provided a fully asynchronous domain decomposed parallel aglorithm as
implemented in production implicit Monte Carlo codes
\citep{brunner_efficient_2009}. We will adapt their algorithm and
directly apply it to a parallel formulation of the Neumann-Ulam
method. Direct analogs can be derived from their works by noting that
the primary difference between solving a linear system with Monte
Carlo methods and fixed source Monte Carlo transport problems is the
content of the Markov chains that are generated. The transitions
represented by these chains are bound by probabilities and weights and
are initiated by the sampling of a discrete source. In the context of
transport problems, those transitions represent events such as
particle scattering and absorption with probabilities that are
determined by physical data in the form of cross sections.

@@

@@

\begin{figure}[htpb!]
  \begin{center}
    \includegraphics[width=6in]{chapters/parallel_mc/titan_pure_strong.pdf}
  \end{center}
  \caption{\textbf{Pure domain decomposition absolute strong scaling.}
    \textit{Super-linear speed-up from memory thrashing in base
      case. MCLS is an order of magnitude slower arithmetically,
      causing the higher efficiency.}}
  \label{fig:titan_pure_strong}
\end{figure}

@@

\begin{figure}[htpb!]
  \begin{center}
    \includegraphics[width=6in]{chapters/parallel_mc/titan_weak_absolute.pdf}
  \end{center}
  \caption{\textbf{Pure domain decomposition absolute weak scaling}}
  \label{fig:titan_weak_absolute}
\end{figure}

\begin{figure}[htpb!]
  \begin{center}
    \includegraphics[width=6in]{chapters/parallel_mc/titan_weak_implementation.pdf}
  \end{center}
  \caption{\textbf{Pure domain decomposition implementation weak
      scaling}}
  \label{fig:titan_weak_implementation}
\end{figure}

@@

@@

\begin{figure}[htpb!]
  \begin{center}
    \includegraphics[width=6in]{chapters/parallel_mc/titan_weak_ms_eff.pdf}
  \end{center}
  \caption{\textbf{Absolute parallel efficiency relative to 16-core
      1-set base case.}}
  \label{fig:titan_weak_ms_eff}
\end{figure}

@@

%%---------------------------------------------------------------------------%%
\section{Multiple Fuel Assembly $SP_N$ Challenge Problem\ }
\label{sec:spn_challenge_problem}

>>>>

%%---------------------------------------------------------------------------%%
\section{Domain Decomposed Neumann-Ulam Algorithm\ }
\label{sec:asynchronous_algorithm}
In the context of radiation transport, in 2009 Brunner and colleagues
provided a fully asynchronous domain decomposed parallel aglorithm as
implemented in production implicit Monte Carlo codes
\citep{brunner_efficient_2009}. We will adapt their algorithm and
directly apply it to a parallel formulation of the Neumann-Ulam
method. Direct analogs can be derived from their works by noting that
the primary difference between solving a linear transport system with
Monte Carlo methods and traditional fixed source Monte Carlo transport
problems is the content of the Markov chains that are generated. The
transitions represented by these chains are bound by probabilities and
weights and are initiated by the sampling of a discrete source. In the
context of transport problems, those transitions represent events such
as particle scattering and absorption with probabilities that are
determined by physical data in the form of cross sections.

@@

\begin{figure}[htpb!]
  \begin{center}
    \includegraphics[width=6in]{chapters/parallel_mc/titan_comm_parameters.pdf}
  \end{center}
  \caption{\textbf{Neumann Ulam parallel algorithm free parameter
      sensitivity.}  \textit{All runtimes were within 6.5\% of the
      fastest observed case.}}
  \label{fig:titan_comm_parameters}
\end{figure}

@@

\begin{figure}[htpb!]
  \begin{center}
    \includegraphics[width=6in]{chapters/parallel_mc/titan_pure_strong.pdf}
  \end{center}
  \caption{\textbf{Pure domain decomposition strong scaling absolute
      efficiency.}  \textit{Super-linear speed-up from memory
      thrashing in base case. MCLS is an order of magnitude slower
      arithmetically, causing the higher efficiency.}}
  \label{fig:titan_pure_strong}
\end{figure}

\begin{figure}[htpb!]
  \begin{center}
    \includegraphics[width=6in]{chapters/parallel_mc/titan_strong_bvsm.pdf}
  \end{center}
  \caption{\textbf{Binary tree vs. master/slave communication scheme
      absolute efficiency.} \textit{Pure domain decomposition absolute
      strong scaling. The master/slave scheme will not scale beyond
      $O(1,000)$ cores due to load imbalance.}}
  \label{fig:titan_strong_bvsm}
\end{figure}

@@

\begin{figure}[htpb!]
  \begin{center}
    \includegraphics[width=6in]{chapters/parallel_mc/titan_weak_absolute.pdf}
  \end{center}
  \caption{\textbf{Pure domain decomposition weak scaling absolute
      efficiency.}}
  \label{fig:titan_weak_absolute}
\end{figure}

\begin{figure}[htpb!]
  \begin{center}
    \includegraphics[width=6in]{chapters/parallel_mc/titan_weak_alg_eff.pdf}
  \end{center}
  \caption{\textbf{Pure domain decomposition weak scaling algorithmic
      efficiency.}}
  \label{fig:titan_weak_algorithmic}
\end{figure}

\begin{figure}[htpb!]
  \begin{center}
    \includegraphics[width=6in]{chapters/parallel_mc/titan_weak_implementation.pdf}
  \end{center}
  \caption{\textbf{Pure domain decomposition weak scaling
      implementation efficiency.}}
  \label{fig:titan_weak_implementation}
\end{figure}

\begin{figure}[htpb!]
  \begin{center}
    \includegraphics[width=6in]{chapters/parallel_mc/titan_weak_bvsm.pdf}
  \end{center}
  \caption{\textbf{Binary tree vs. master/slave communication scheme
      absolute efficiency.}  \textit{Pure domain decomposition
      absolute weak scaling. The master/slave scheme will not scale
      beyond $O(1,000)$ cores due to load imbalance.}}
  \label{fig:titan_weak_bvsm}
\end{figure}

@@

\begin{table}[h!]
  \begin{center}
    \begin{tabular}{lcc}\hline\hline
      \multicolumn{1}{c}{Case}& 
      \multicolumn{1}{c}{Iterations}&
      \multicolumn{1}{c}{$\eta_{alg}$} \\\hline
      1 & 22 & 1.0 \\
      2-split & 22 & 1.0 \\
      2-replicated & 16 & 1.375 \\
      4-split & 22 & 1.0 \\
      4-replicated & 13 & 1.692 \\
      %%
      \hline\hline
    \end{tabular}
  \end{center}
  \caption{\textbf{Strong scaling algorithmic efficiency using
      multiple sets.}}
  \label{tab:ms_strong_alg_eff}
\end{table}

\begin{figure}[htpb!]
  \begin{center}
    \includegraphics[width=6in]{chapters/parallel_mc/titan_strong_ms_impeff.pdf}
  \end{center}
  \caption{\textbf{Strong scaling implementation efficiency relative
      to 16-core 1-set base case.}}
  \label{fig:titan_strong_ms_impeff}
\end{figure}

@@

For a multiple set analysis of the weak scaling results, we have to
consider that adding sets to the problem is actually a form of strong
scaling. Consider a single set problem ($S=1$) for a given $P$ and
$N$. When that problem is now solved using 2 sets we have $S=2$, $N$
and $2P$. For any number of sets we then have a global problem size of
$N$ while the number of processors is $SP$ where $P$ is the number of
processors in the single set case. Therefore, $N/P$ is no longer fixed
in the weak scaling study but rather $N/(SP)$ is fixed. Therefore, we
modify the weak scaling absolute efficiency relationship to account
for this fact:
\begin{equation}
\eta_{weak}(M|N,P|Q,S) = \frac{T(M,Q)}{S \times T(N,P)}\:,
  \label{eq:ms_weak_efficiency}
\end{equation}
where now the absolute efficiency is a function of the number of sets
in the problem. For problems where adding sets improves time to
solution, this new weak scaling formulation accounts for how
efficiently those extra resources are incorporated into the
problem. Without such a modification, the largest case run for the
split 4 set case would have a computed weak scaling efficiency of
nearly 100\% using the standard formula for absolute
efficiency. However, using 4 times as many cores as the single set
case increases the runtime for a fixed global problem size and
therefore we actually expect to compute a very low efficiency that
shows this very poor use of resources.

\begin{figure}[htpb!]
  \begin{center}
    \includegraphics[width=6in]{chapters/parallel_mc/titan_weak_ms_eff.pdf}
  \end{center}
  \caption{\textbf{Weak scaling absolute efficiency relative to
      16-core 1-set base case.}}
  \label{fig:titan_weak_ms_eff}
\end{figure}

\begin{table}[h!]
  \begin{center}
    \begin{tabular}{lcc}\hline\hline
      \multicolumn{1}{c}{Case}& 
      \multicolumn{1}{c}{Iterations}&
      \multicolumn{1}{c}{$\eta_{alg}$} \\\hline
      1 & 22 & 1.0 \\
      2-split & 22 & 1.0 \\
      2-replicated & 16 & 1.375 \\
      4-split & 22 & 1.0 \\
      4-replicated & 13 & 1.692 \\
      %%
      \hline\hline
    \end{tabular}
  \end{center}
  \caption{\textbf{Weak scaling algorithmic efficiency using multiple
      sets.}}
  \label{tab:ms_weak_alg_eff}
\end{table}

\begin{figure}[htpb!]
  \begin{center}
    \includegraphics[width=6in]{chapters/parallel_mc/titan_weak_ms_impeff.pdf}
  \end{center}
  \caption{\textbf{Implementation parallel efficiency relative to 16-core
      1-set base case.}}
  \label{fig:titan_weak_ms_impeff}
\end{figure}

@@
